\documentclass{article}
\usepackage{hyperref}

\title{Rigidity vs. Harmony: Themes Across Star Wars, Ovid's \textit{Metamorphoses}, and Kid Cudi's Music}
\author{}
\date{}

\begin{document}

\maketitle

The Star Wars saga, with its rich narrative and philosophical depth, offers a profound exploration of the themes of rigidity and dominance, particularly through the contrasting philosophies of the Sith and the Jedi. These themes are not confined to the realm of science fiction but resonate across classical literature and modern music, such as the works of Ovid and Kid Cudi. This essay delves into these themes, examining their manifestations and implications across different mediums.

\section*{Sith Rigidity and Dominance in Star Wars}

The Sith, as depicted in Star Wars, embody a rigid and uncompromising nature, driven by an insatiable desire for power and control. This is encapsulated in their philosophy and actions:

\begin{itemize}
    \item \textbf{Sith Code}: The Sith Code emphasizes strength, power, and victory, often at the expense of others. It promotes a worldview where the strong dominate the weak, and emotions like anger and hatred are harnessed to gain power. This philosophy is inherently rigid, leaving little room for compassion or flexibility \cite{serter2021light}.
    \item \textbf{Rule of Two}: Established by Darth Bane, the Rule of Two dictates that there should only be a master and an apprentice. This rule creates a cycle of betrayal, as the apprentice is expected to eventually overthrow the master. The Rule of Two exemplifies the Sith's uncompromising pursuit of power, where loyalty is transient and power is the ultimate goal \cite{forbes1999battling}.
    \item \textbf{Key Characters}: Darth Sidious (Emperor Palpatine) and Darth Vader are quintessential representations of the Sith's rigid nature. Palpatine's meticulous manipulation of the Galactic Senate and the Jedi Order, coupled with Vader's ruthless enforcement of the Emperor's will, highlight their relentless and uncompromising pursuit of dominance \cite{springer2019star}.
\end{itemize}

\section*{Jedi Balance and Harmony}

In stark contrast, the Jedi strive for balance and harmony, both within themselves and the galaxy:

\begin{itemize}
    \item \textbf{Jedi Code}: The Jedi Code emphasizes self-discipline, compassion, and the greater good. Jedi seek to understand and control their emotions, promoting peace and harmony. This philosophy is inherently flexible, allowing for growth and adaptation \cite{serter2021light}.
    \item \textbf{Selflessness and Service}: Jedi are trained to be selfless and to serve others, acting as guardians of peace and justice. Their actions are guided by a sense of duty and compassion, in stark contrast to the Sith's self-serving ambitions \cite{forbes1999battling}.
    \item \textbf{Key Characters}: Obi-Wan Kenobi and Luke Skywalker embody the Jedi's values. Obi-Wan's compassion and mentorship, and Luke's belief in redemption and forgiveness, highlight the Jedi's commitment to balance and selflessness. Their actions are driven by a desire to protect and uplift others, rather than dominate them \cite{springer2019star}.
\end{itemize}

\section*{Ovid's \textit{Metamorphoses}}

Ovid's \textit{Metamorphoses} is a rich tapestry of myths and legends, exploring themes of transformation, power, and the human condition. The work is filled with examples of rigidity and dominance, often depicted through the actions of gods and mortals:

\begin{itemize}
    \item \textbf{Transformation and Power}: The gods in \textit{Metamorphoses} often transform mortals as a form of punishment or reward. These transformations can be literal, such as when Diana turns Actaeon into a deer, or metaphorical, highlighting an essential quality of the transformed person. For example, Jupiter turns Lycaon into a wolf, reflecting Lycaon's bloodthirsty nature \cite{sharrock2020metamorphic}.
    \item \textbf{Exercise of Divine Power}: The gods' ability to transform mortals underscores their dominance and control over human fate. This theme of power and dominance is central to many stories in \textit{Metamorphoses}, illustrating the often arbitrary and rigid nature of divine authority \cite{gallagher2014metamorphosis}.
    \item \textbf{Humanity vs. Nature}: Ovid portrays humanity's attempt to wield power over nature, reflecting the unique position of humans as creatures capable of dominion. This theme is evident in stories where mortals challenge natural order or divine will, often leading to their downfall. The story of Phaethon, who attempts to drive the chariot of the sun, illustrates the catastrophic consequences of overreaching ambition and the rigid boundaries set by nature and the gods \cite{sharrock2020metamorphic}.
    \item \textbf{Love and Destruction}: Love in \textit{Metamorphoses} is often portrayed as a destructive force, leading characters to betray their families, kingdoms, or values. This theme aligns with the Sith's embrace of powerful emotions and the resulting chaos. The tragic story of Myrrha, who falls in love with her father and betrays the laws of nature, exemplifies the destructive power of forbidden love. Similarly, the tale of Scylla, who betrays her father and kingdom for Minos, highlights the devastating impact of passion \cite{gallagher2014metamorphosis}.
\end{itemize}

\section*{Kid Cudi's Music}

Kid Cudi's music often explores themes of mental health, self-discovery, and the struggle between darkness and light, aligning with both Jedi and Sith philosophies:

\begin{itemize}
    \item \textbf{Mental Health and Struggle}: Songs like \textit{Pursuit of Happiness} and \textit{Soundtrack 2 My Life} address Cudi's battles with depression and anxiety, resonating with the Sith's embrace of powerful emotions. Cudi's openness about his struggles highlights the importance of mental health awareness and the complexities of human emotions \cite{aditi2025insights,holr2025close}.
    \item \textbf{Self-Discovery and Growth}: Tracks like \textit{Day 'n' Nite} and \textit{Reborn} reflect Cudi's journey of self-discovery and personal growth, akin to the Jedi's pursuit of balance and harmony. His music often emphasizes resilience and the importance of understanding oneself \cite{aditi2025insights,holr2025close}.
    \item \textbf{Dreams and Aspirations}: Songs such as \textit{Enter Galactic (Love Connection Part I)} and \textit{Up Up \& Away} emphasize the importance of pursuing dreams despite obstacles, embodying hope and determination. This aligns with the Jedi's focus on selflessness and the greater good, as well as the Sith's drive for personal ambition \cite{aditi2025insights,holr2025close}.
\end{itemize}

\section*{Conclusion}

The themes of rigidity and dominance versus balance and harmony are timeless and universal, resonating across different mediums and eras. From the epic battles of Star Wars to the tragic conflicts in Ovid's \textit{Metamorphoses} and the introspective journeys in Kid Cudi's music, these themes continue to captivate and inspire audiences. They remind us of the importance of balance, self-awareness, and the choices that define our paths.

This literary critique underscores the enduring relevance of these themes, inviting readers to reflect on their own lives and the world around them. Whether through the lens of a galaxy far, far away, the pages of a classic myth, or the beats of a modern song, the exploration of rigidity and dominance versus balance and harmony offers profound insights into the human condition.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
